Guaranteeing Safety and Liveness of Unmanned Aerial Vehicle Platoons on Air Highways

### 2016-02-13 docs/AIAA 2016/high_sim.tex
no message
Updated AIAA 2016 paper

\subsubsection{Cost-Minimizing Paths}
Fig. \ref{fig:airHighway_results} shows the San Francisco Bay Area (geographic) map, cost map, cost-minimizing paths, and contours of the value function. The region enclosed by the black boundary represent ``regoin around airports'', which have the highest cost. The dark blue, yellow, and light blue regions represent the ``cities'', the ``water'', and the ``other''' categories, respectively. We assumed that the origin corresponds to the city ``Concord'', and chose a number of other major cities as destinations.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.75\textwidth]{"fig/airHighway_results"}
	\caption{Cost-minimizing paths computed by the Fast Marching Method based on the assumed cost map of the San Francisco Bay Area}
	\label{fig:airHighway_results}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.75\textwidth]{"fig/airHighway_sparse"}
	\caption{Results of conversion from cost-minimizing paths to highway way points.}
	\label{fig:airHighway_sparse}
\end{figure}

>>>>

\subsubsection{Cost-Minimizing Paths}
Figure \ref{fig:airHighway_results} shows the San Francisco Bay Area (geographic) map, cost map, cost-minimizing paths, and contours of the value function. The region enclosed by the black boundary represent ``regoin around airports'', which have the highest cost. The dark blue, yellow, and light blue regions represent the ``cities'', the ``water'', and the ``other''' categories, respectively. We assumed that the origin corresponds to the city ``Concord'', and chose a number of other major cities as destinations.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{"fig/airHighway_results"}
	\caption{Cost-minimizing paths computed by the Fast Marching Method based on the assumed cost map of the San Francisco Bay Area}
	\label{fig:airHighway_results}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{"fig/airHighway_sparse"}
	\caption{Results of conversion from cost-minimizing paths to highway way points.}
	\label{fig:airHighway_sparse}
\end{figure}

### 2016-02-13 docs/AIAA 2016/introduction.tex
Updated AIAA 2016 paper

% !TEX root = main.tex
\section{Introduction}
Unmanned aerial vehicle (UAV) systems have in the past been mainly used for military operations \cite{Tice91}. Recently, however, there has been an immense surge of interest in using UAVs for civil applications. Through projects such as Amazon Prime Air and Google Project Wing, companies are looking to send UAVs into the airspace to deliver packages \cite{PrimeAir,ProjectWing}. As a rough estimate, suppose in a city of 2 million people, each person requests a drone delivery every 2 months on average and each delivery requires a 30 minute trip for the drone. This would equate to thousands of vehicles simultaneously in the air. As a result, government agencies such as the Federal Aviation Administration (FAA) and National Aeronautics and Administration (NASA) of the United States are also investigating air traffic control for autonomous vehicles in order to prevent collisions among potentially numerous UAVs \cite{FAA13}. Applications of UAVs extend beyond package delivery; they can also be used, for example, to provide supplies or to firefight in areas that are difficult to reach but require prompt response \cite{Debusk10}.

Optimal control and game theory are powerful tools for providing liveness and safety guarantees to controlled dynamical systems under disturbances, and various formulations \cite{Bokanowski10,Mitchell05,Barron89} have been successfully used to analyze problems involving a small number of vehicles \cite{Fisac15,Chen14,Ding08}. These formulations are based on Hamilton-Jacobi (HJ) reachability, which computes the reachable set, defined as the set of states from which a system is guaranteed to have a control strategy to reach a target set of states. Reachability is a powerful tool because reachable sets can be used for synthesizing both controllers that steer the system towards a set of goal states (liveness controllers), and controllers that steer the system away from a set of unsafe states (safety controllers). Furthermore, the HJ formulations are flexible in terms of system dynamics, enabling the analysis of non-linear systems. The power and success of HJ reachability analysis in previous applications cannot be denied, especially since numerical tools are readily available to solve the associated HJ Partial Differential Equation (PDE) \cite{LSToolbox,Osher02,Sethian96}. However, the computation is done on a grid, making the problem complexity scale exponentially with the number of states, and therefore with the number of vehicles. This makes the computation intractable for large numbers of vehicles.

@@

Previous analysis of a large number of vehicles typically do not provide liveness and safety guarantees to the extent that HJ reachability does; however, HJ reachability typically cannot be used to tractably analyze a large number of vehicles. %In the context of HJ reachability, putting vehicles into platoons is desirable because of the additional structure that platoons impose on its members. With additional structure, pairwise safety guarantees of vehicles can be more easily translated into safety guarantees of all the vehicles in the platoon. 
In this paper, we attempt to reconciliate this trade-off by assuming a single-file platoon, which provides structure that allows pairwise safety guarantees from HJ reachability to translate to safety guarantees for the whole platoon. We first propose a hybrid systems model of UAVs in platoons to capture this structure. Then, we show how HJ reachability can be used to synthesize \textit{liveness controllers} that enable vehicles to reach a set of desired states, and wrap \textit{safety controllers} around the liveness controllers in order to prevent dangerous configurations such as collisions. Finally, we show simulation results of quadrotors forming a platoon, a platoon responding to a malfunctioning member, and a platoon responding to an outside intruder to illustrate the behavior of vehicles in these scenarios and demonstrate the guarantees provided by HJ reachability.

%Motivation:
%\begin{itemize}
%\item Applications of UAVs, potential numbeurs
%\item Importance of safety guarantees
%\item Computation complexity
%\end{itemize}
%
%Related work:
%\begin{itemize}
%\item Platooning references: lack of (?) safety guarantees
%\begin{itemize}\item\textcolor{blue}{Mention PATH papers \& Lygeros thesis}\end{itemize}
%\item HJI, safety guarantees, limited by dimensionality
%\end{itemize}
%
%Summary of results
%\begin{itemize}
%\item Hybrid systems model of vehicles
%\item Reachability guarantees wrapped around existing methods
%\item Reachability offers flexibility in terms of design
%\item Illustrative platoon functions
%\end{itemize}

>>>>

% !TEX root = main.tex
\section{Introduction}
Unmanned aerial vehicle (UAV) systems have in the past been mainly used for military operations \cite{Tice91}. Recently, however, there has been an immense surge of interest in using UAVs for civil applications. Through projects such as Amazon Prime Air \cite{PrimeAir} and Google Project Wing \cite{ProjectWing}, companies are looking to send UAVs into the airspace to not only deliver commerical packages, but also for important tasks such as aerial surveillance, emergency supply delivery, and search and rescue. In the future, the applications of UAVs are only limited by human imagination.

As a rough estimate, suppose in a city of 2 million people, each person requests a drone delivery every 2 months on average and each delivery requires a 30 minute trip for the drone. This would equate to thousands of vehicles simultaneously in the air just from package delivery services. Applications of UAVs extend beyond package delivery; they can also be used, for example, to provide supplies or to firefight in areas that are difficult to reach but require prompt response \cite{Debusk10}. As a result, government agencies such as the Federal Aviation Administration (FAA) and National Aeronautics and Administration (NASA) are also investigating air traffic control for autonomous vehicles in order to prevent collisions among potentially numerous UAVs \cite{FAA13, NASA16}.

Optimal control and game theory are powerful tools for providing liveness and safety guarantees to controlled dynamical systems under bounded disturbances, and various formulations \cite{Bokanowski10,Mitchell05,Barron89} have been successfully used to analyze problems involving a small number of vehicles \cite{Fisac15,Chen14,Ding08}. These formulations are based on Hamilton-Jacobi (HJ) reachability, which computes the reachable set, defined as the set of states from which a system is guaranteed to have a control strategy to reach a target set of states. Reachability is a powerful tool because reachable sets can be used for synthesizing both controllers that steer the system towards a set of goal states (liveness controllers), and controllers that steer the system away from a set of unsafe states (safety controllers). Furthermore, the HJ formulations are flexible in terms of system dynamics, enabling the analysis of non-linear systems. The power and success of HJ reachability analysis in previous applications cannot be denied, especially since numerical tools are readily available to solve the associated HJ Partial Differential Equation (PDE) \cite{LSToolbox,Osher02,Sethian96}. However, the computation is done on a grid, making the problem complexity scale exponentially with the number of states, and therefore with the number of vehicles. This makes the computation intractable for large numbers of vehicles.

In order to accommodate potentially thousands of vehicles simultaneously flying in the air, additional structure is needed to allow for tractable analysis and intuitive monitoring. An air highway system on which platoons of vehicles travel accomplishes both goals. However, many details of such a concept need to be addressed. Due to the flexibility of placing air highways compared to building ground highways in terms of highway location, even the problem of air highway placement can be daunting task. To address this, in the first part of this paper, we propose a flexible and computationally efficient way to perform optimal air highway replacement given an arbitrary cost map that can capture the desirability of having UAVs fly over any geographical location. We demonstrate our method using the San Francisco Bay Area as an example. Once air highways are in place, platoons of UAVs can then fly in fixed formations along the highway to get from origin to destination.

@@

Previous analysis of a large number of vehicles typically do not provide liveness and safety guarantees to the extent that HJ reachability does; however, HJ reachability typically cannot be used to tractably analyze a large number of vehicles. In the second part of this paper, we propose organizing UAVs into platoons, which provide structure that allows pairwise safety guarantees from HJ reachability to translate to safety guarantees for the whole platoon. With respect to platooning, we first propose a hybrid systems model of UAVs in platoons to establish the modes of operation needed for our platooning concept. Then, we show how reachability-based controllers can be synthesized to enable vehicles successfully perform mode switching, as well as prevent dangerous configurations such as collisions. Finally, we show several simulations to illustrate the behavior of vehicles in various scenarios and demonstrate the guarantees provided by HJ reachability.

### 2016-02-13 docs/AIAA 2016/plat_controllers.tex
Updated AIAA 2016 paper

% !TEX root = main.tex
\subsection{Reachability-Based Controllers}
Reachability analysis is useful for constructing controllers in a large variety of situations. In order to construct different controllers, an appropriate target set needs to be defined depending on the goal of the controller. If one defines the target set to be a set of desired states, the reachable set, once computed, would represent the states from which a system needs to first arrive at in order to reach the desired states. On the other hand, if the target set represents a set of undesirable states, then the reachable set would indicate the region of the state space that the system needs to avoid. In addition, the system dynamics with which the reachable set is computed provides additional flexibility when using reachability to construct controllers.

@@

\subsubsection{Getting to a Target State}
The controller used by a vehicle to a target state is important in a couple of situations in the platooning context. First, a vehicle in the ``Free'' mode can use the controller to merge onto a highway, forming a platoon and changing modes to a ``Leader'' vehicle. Second, a vehicle in either the ``Leader'' mode or the ``Follower'' mode can use this controller to change to a different highway, changing modes to a ``Leader'' vehicle.

@@

\subsubsection{Getting to a State Relative to Another Vehicle}
In the platooning context, being able to go to a state relative to another moving vehicle is important for the purpose of forming and joining platoons. For example, a ``Free'' vehicle may join an existing platoon that is on a highway and change modes to become a ``Follower''. Also, a ``Leader'' or ``Follower'' may join another platoon and afterwards go into the ``Follower'' mode.

@@

\subsubsection{Avoiding Collision}
A vehicle can use a liveness controller described in the previous sections when it is not in any danger of collision with other vehicles. If the vehicle could potentially be involved in a collision within the next short period of time, it must switch to a safety controller. The safety controller is available in every mode, and executing the safety controller to perform an avoidance maneuver does not change a vehicle's mode.

@@

THE FOLLOWING SHOULD BE IN ANOTHER SECTION

\subsubsection{Other Quadrotor Maneuvers}
Reachability was used in Sections \ref{subsec:highway_merge} and \ref{subsec:platoon_merge} for the relatively complex maneuvers. For the simpler maneuvers of traveling along a highway and following a platoon, we resort to simpler controllers described below.

>>>>

% !TEX root = main.tex
\subsection{Reachability-Based Controllers \label{sec:reach_ctrl}}
Reachability analysis is useful for constructing controllers in a large variety of situations. In order to construct different controllers, an appropriate target set needs to be defined depending on the goal of the controller. If one defines the target set to be a set of desired states, the reachable set, once computed, would represent the states from which a system needs to first arrive at in order to reach the desired states. On the other hand, if the target set represents a set of undesirable states, then the reachable set would indicate the region of the state space that the system needs to avoid. In addition, the system dynamics with which the reachable set is computed provides additional flexibility when using reachability to construct controllers.

@@

\subsubsection{Getting to a Target State \label{sec:abs_target_ctrl}}
The controller used by a vehicle to a target state is important in a couple of situations in the platooning context. First, a vehicle in the ``Free'' mode can use the controller to merge onto a highway, forming a platoon and changing modes to a ``Leader'' vehicle. Second, a vehicle in either the ``Leader'' mode or the ``Follower'' mode can use this controller to change to a different highway, changing modes to a ``Leader'' vehicle.

@@

\subsubsection{Getting to a State Relative to Another Vehicle \label{sec:rel_target_ctrl}}
In the platooning context, being able to go to a state relative to another moving vehicle is important for the purpose of forming and joining platoons. For example, a ``Free'' vehicle may join an existing platoon that is on a highway and change modes to become a ``Follower''. Also, a ``Leader'' or ``Follower'' may join another platoon and afterwards go into the ``Follower'' mode.

@@

\subsubsection{Avoiding Collisions \label{sec:collision_ctrl}}
A vehicle can use a liveness controller described in the previous sections when it is not in any danger of collision with other vehicles. If the vehicle could potentially be involved in a collision within the next short period of time, it must switch to a safety controller. The safety controller is available in every mode, and executing the safety controller to perform an avoidance maneuver does not change a vehicle's mode.

@@

\subsection{Other Quadrotor Maneuvers \label{sec:other_ctrl}}
Reachability was used in Section \ref{sec:reach_ctrl} for the relatively complex maneuvers that require safety and liveness guarantees. For the simpler maneuvers of traveling along a highway and following a platoon, many well-known classical controllers suffice. For illustration, we use the simple controllers described below.

### 2016-02-13 docs/AIAA 2016/plat_form.tex
Updated AIAA 2016 paper

If a vehicle from any of the other modes malfunction, it transitions into the Faulty mode and descends after a duration of $t_\text{internal}$.
\end{itemize}

@@

Suppose that there are $N$ vehicles in total. We consider a platoon of vehicles to be a group of $M\le N$ vehicles, denoted $Q_{P_1}, \ldots, Q_{P_M}, \{P_j\}_{j=1}^M \subseteq \{i\}_{i=1}^N$, in a single-file formation. We will assume that the vehicles in a platoon travel along an air highway. The vehicles maintain a separation distance of $b$ with its neighbors inside the platoon. In order to allow for close proximity of the vehicles and the ability to resolve multiple simultaneous safety breaches, we assume that in the event of a malfunction, a vehicle will be able to exit the altitude range of the highway within a duration of $t_\text{internal}=1.5$. Such a requirement may be implemented practically as an emergency landing procedure to which the vehicles revert when a malfunction is detected.

>>>>

If a vehicle from any of the other modes malfunction, it transitions into the Faulty mode and descends after some pre-specified duration $t_\text{faulty}$.
\end{itemize}

@@

Suppose that there are $N$ vehicles in total. We consider a platoon of vehicles to be a group of $M\le N$ vehicles, denoted $Q_{P_1}, \ldots, Q_{P_M}, \{P_j\}_{j=1}^M \subseteq \{i\}_{i=1}^N$, in a single-file formation. We will assume that the vehicles in a platoon travel along an air highway. The vehicles maintain a separation distance of $b$ with its neighbors inside the platoon. In order to allow for close proximity of the vehicles and the ability to resolve multiple simultaneous safety breaches, we assume that in the event of a malfunction, a vehicle will be able to exit the altitude range of the highway within a duration of $t_\text{faulty}=1.5$. Such a requirement may be implemented practically as an emergency landing procedure to which the vehicles revert when a malfunction is detected.
